Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf

\subsection{LD Sequence Generators} A lattice, digital sequence, or Halton sequence generator typically supplies the sequence $\bsx_1, \bsx_2, \ldots$.  These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{latticepts.eps}
    \caption{A two-dimensional projection of}
    \label{fig:latticepts}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five. These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.  }
    \label{fig:beam}
\end{figure}

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.  We illustrate in Figure \ref{fig:beam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.  The beam position is discretized at $31$ points.  QMCPy computes the expected deflection of the beam at these points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

@@

\subsection{Stopping Criteria}
The cantilevered beam example illustrated in Figure \ref{fig:beam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.  See \cite{HicJim16a} for details of this stopping criterion for lattice rules and \cite{HicEtal18a} for a review on automatic QMC stopping criteria.  Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.  Some of these stopping criteria for absolute error criteria, \eqref{CDHJS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.

@@

A fully coded example of this problem is available at \url{https://github.com/QMCSoftware/QMCSoftware/blob/master/demos/umbridge.ipynb}. At a high level, one begins by running the UM-Bridge model in a Docker container. A model is then initialized in the UM-Bridge Python client which is then provided to QMCPy alongside a low discrepancy point generator and distribution governing the $3$ beam stiffness values. QMCPy then performs adaptive QMC cubature to simultaneously approximate the expected displacement at the $31$ locations of interest. When the cubature method queries the model, the QMCPy wrapper around UM-Bridge model sends requests to the Docker UM-Bridge model in parallel. The load balancing provided by UM-Bridge may be utilized to evaluate the model in parallel, a potentially advantageous feature for models which are costly to evaluate. Figure \ref{fig:umbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.

\begin{figure}
    \centering
    \includegraphics[height=4cm]{umbridge_muq.png}
    \caption{QMCPy approximate expected displacement at $31$ locations subject to random uniform stiffness. }
    \label{fig:umbridge_muq}
\end{figure}

>>>>

\subsection{LD Sequence Generators} A lattice, digital sequence, or Halton sequence generator typically supplies the sequence $\bsx_1, \bsx_2, \ldots$.  These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.  Figure \ref{CDHJSfiglatticepts} illustrates an example of a lattice sequence.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{latticepts.eps}
    \caption{A two-dimensional projection of lattice points for increasing sample size.  The later points fill in gaps because they are highly correlated with the earlier points.  These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}. }
    \label{CDHJSfiglatticepts}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five. These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.  }
    \label{CDHJSfigbeam}
\end{figure}

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{umbridge}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$.  We illustrate in Figure \ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam with random Young's modulus \cite{ParSee22a}, where the randomness is given by $\bsX \sim \calU[1,1.2]^3$.  The beam position is discretized at $31$ points.  QMCPy computes the expected deflection of the beam at these points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

@@

\subsection{Stopping Criteria}
The cantilevered beam example illustrated in Figure \ref{CDHJSfigbeam} utilizes QMCPy's stopping criterion to choose the sample size, $n$, based on the input error tolerance, $\varepsilon$, and the observed discrete Fourier series coefficients of the integrand, $f$.  See \cite{HicJim16a} for details of this stopping criterion for lattice rules and \cite{HicEtal18a} for a review on automatic QMC stopping criteria.  Theoretically justified stopping criteria for choosing sample sizes are important and a relatively recent development \cite{HicEtal14a,HicJim16a,JimHic16a,RatHic19a,JagHic22a}.  Some of these stopping criteria for absolute error criteria, \eqref{CDHJS_eq:error_tol},  have been extended to relative error criteria or a combination of absolute and relative error criteria, as well as stopping criteria for approximating functions of several integrals \cite{Jia16a,GilJim16b,HicEtal17a,JagSor23a}.

@@

A fully coded example of this problem is available at \url{https://github.com/QMCSoftware/QMCSoftware/blob/master/demos/umbridge.ipynb}. At a high level, one begins by running the UM-Bridge model in a Docker container. A model is then initialized in the UM-Bridge Python client which is then provided to QMCPy alongside a low discrepancy point generator and distribution governing the $3$ beam stiffness values. QMCPy then performs adaptive QMC cubature to simultaneously approximate the expected displacement at the $31$ locations of interest. When the cubature method queries the model, the QMCPy wrapper around UM-Bridge model sends requests to the Docker UM-Bridge model in parallel. The load balancing provided by UM-Bridge may be utilized to evaluate the model in parallel, a potentially advantageous feature for models which are costly to evaluate. Figure \ref{CDHJSfigumbridge_muq} depicts the $31$ approximate expected displacement values subject to uniform stiffness.

\begin{figure}
    \centering
    \includegraphics[height=4cm]{umbridge_muq.png}
    \caption{QMCPy approximate expected displacement at $31$ locations subject to random uniform stiffness. }
    \label{CDHJSfigumbridge_muq}
\end{figure}
